\Large\textbf{General Stage Untangling with Applications to Algorithm Derivation}

### 2014-02-21 paper/paper.tex
Worked out most of the proof for the if case.

\begin{figure}
\caption{\corelang~Reference Semantics}
\label{fig:refSemantics}
\begin{mathpar}
\inferRedsplitone [\rmunit]		{\red {()}{()}{()}}												{\cdot} 												\and
\inferRedsplitone [int]			{\red {i}{i}{()}}													{\cdot} 												\and
\inferRedsplitone [bool]			{\red {b}{b}{()}}													{\cdot} 												\\
\inferSpeculation [\rmunit]		{\red {()}{()}}													{\cdot} 												\and
\inferSpeculation [int]			{\red {i}{i}}														{\cdot} 												\and
\inferSpeculation [bool]			{\red {b}{b}}														{\cdot} 												\\
\inferRedsplitone [let]			{\red {\letin{x}{e_1}{e_2}}{u_2}{\letin {x}{r_1}{r_2}}}		{\sub [1] & \red{[[u_1,x]/x]e_2}{u_2}{r_2}} 		\and
\inferSpeculation [let]			{\red {\letin{x}{e_1}{e_2}}{\letin{x}{r_1}{r_2}}}				{\sub [1] & \sub [2]}									\\
\inferRedsplitone [hyp]			{\red {[u,x]}{u}{x}}												{\cdot}													\and
\inferRedsplitone [\times I]		{\red {(e_1,e_2)}{\valprod{u_1}{u_2}}{\valprod{r_1}{r_2}}}	{\sub [1] & \sub [2]}									\and
\inferSpeculation [\times I]		{\red {(e_1,e_2)}{(r_1,r_2)}}									{\sub [1] & \sub [2]}									\and
\inferRedsplitone [\times E_1]	{\red {\pi_1~e}{u_1}{\pi_1 r}}									{\red {e}{\valprod{u_1}{u_2}}{r}}					\and
\inferSpeculation [\times E_1]	{\red {\pi_1~e}{\pi_1 r}}										{\red {e}{r}}											\and
\inferRedsplitone [\times E_2]	{\red {\pi_2~e}{u_2}{\pi_2 r}}									{\red {e}{\valprod{u_1}{u_2}}{r}} 					\and
\inferSpeculation [\times E_2]	{\red {\pi_2~e}{\pi_2 r}}										{\red {e}{r}} 											\and
\inferRedsplitone [if_T] 		{\red {\ifthen{e_1}{e_2}{e_3}}{u}{r_2}}							{\red{e_1}{\bf true }{r_1} & \red{e_2}{u}{r_2}}	\and
\inferRedsplitone [if_F] 		{\red {\ifthen{e_1}{e_2}{e_3}}{u}{r_3}}							{\red{e_1}{\bf false}{r_1} & \red{e_3}{u}{r_3}}	\\
\inferSpeculation [if] 			{\red {\ifthen{e_1}{e_2}{e_3}}{\ifthen{r_1}{r_2}{r_3}}}		{\sub [1] & \sub [2] & \sub [3]}					\\
\inferRedsplitone [\fut I]		{\red {\next~e}{()}{r}}											{\rstwosub}											\and
\inferSpeculation [\fut E]		{\red {\prev~e}{r}}												{\rsone {e}{()}{r}}
\end{mathpar}
\end{figure}

@@

\begin{figure*}
\caption{Term Splitting}
\label{fig:termSplit}
\newcommand {\projbind} [1] {\letin{l_{#1}}{\pi_{#1} l}{r_{#1}}}
\begin{mathpar}
\infersplitone [unit]			{\spl {()}{\rmunit}{((),()),\_.()}}															{\cdot} \and
\infersplittwo [unit]			{\spl {()}{\rmunit}{(),\_.()}}																{\cdot} \and
\infersplitone [int]			{\spl {i}{\rm int}{(i,()),\_.()}}															{\cdot} \and
\infersplittwo [int]			{\spl {i}{\rm int}{(),\_.i}}																	{\cdot} \and
\infersplitone [\times I] 	{\spl {(e_1,e_2)}{A\times B}
									{ \begin{array}{l}
									\left(
										\letin{y_1}{c_1}{\letin{y_2}{c_2}{
										((\pi_1 y_1, \pi_1 y_2),(\pi_2 y_1, \pi_2 y_2))}}
									\right), \\
									l.(\projbind 1,\projbind 2)
									\end{array}}}																				{\sub 1 A & \sub 2 B} \and
\infersplittwo [\times I]	{\spl {(e_1,e_2)}{A\times B}{(p_1,p_2),l.(\projbind 1,\projbind 2)}}					{\sub 1 A & \sub 2 B} \and
\infersplitone [\times E_1]	{\spl {\pi_1~e}{A}{\letin{y}{c}{(\pi_1(\pi_1 y),\pi_2 y)},l.\pi_1~r}}					{\sub {} {A\times B}} \and
\infersplitone [\times E_2]	{\spl {\pi_2~e}{B}{\letin{y}{c}{(\pi_2(\pi_1 y),\pi_2 y)},l.\pi_2~r}}					{\sub {} {A\times B}} \and
\infersplittwo [\times E_1]	{\spl {\pi_1~e}{A}{p,l.\pi_1~r}}															{\sub {} {A\times B}} \and
\infersplittwo [\times E_2]	{\spl {\pi_2~e}{B}{p,l.\pi_2~r}}															{\sub {} {A\times B}} \and
\infersplitone [hyp]			{\spl {v}{A}{(x,()),\_.()}}																	{x:A^\bbone \in \Gamma} \and
\infersplittwo [hyp]			{\spl {v}{A}{(),\_.x}}																		{x:A^\bbtwo \in \Gamma} \\
\infersplitone [let]			{\spl {\letin{x}{e_1}{e_2}}{B}{\cdots}}														{\sub 1 A & \sub [\Gamma,x:A^\bbone] 2 B} \and
\infersplittwo [let]			{\spl {\letin{x}{e_1}{e_2}}{B}{(p_1,p_2), l.\letin{x}{(\projbind 1)}{\projbind 2}}}	{\sub 1 A & \sub [\Gamma,x:A^\bbtwo] 2 B} \and
\infersplitone [\fut I]		{\spl {\next~e}{\fut A}{((),p),l.r}}														{\splittwosub {} A} \and
\infersplittwo [\fut E]		{\spl {\prev~e}{A} {\pi_2~c,l.r}}															{\splitonesub {} {\fut A}} \and
\infersplitone [if]			{\spl {\left( \tallif {e_1}{e_2}{e_3} \right)}{A} 
								{\begin{array}{l}
									\left(\tallif {\pi_1 c_1}
										{\letin{y}{c_2}{(\pi_1 y, \inl(\pi_2 y))}}
										{\letin{y}{c_3}{(\pi_1 y, \inr(\pi_2 y))}}
									\right), \\
									l.(\caseof{\pi_1 l}{l_2.r_2}{l_3.r_3})
								\end{array}}}																					{\sub 1 \rmbool & \sub 2 A & \sub 3 A } \and
\infersplittwo [if]			{\spl {\left(\tallif {e_1}{e_2}{e_3}\right)}{A} 
								{(p_1,p_2,p_3),l.\left(\tallif{(\projbind 1)}{(\projbind 2)}{(\projbind 3)}\right)}}	{\sub 1 \rmbool & \sub 2 A & \sub 3 A } \and
\end{mathpar}
\end{figure*}

@@

\item {\bf If.} Suppose that
\begin{enumerate}
\item $\splitone {\left( \tallif {e_1}{e_2}{e_3} \right)}{A} 
								{\begin{array}{l}
									\left(\tallif {\pi_1c_1}
										{\letin{y}{c_2}{(\pi_1 y, \inl(\pi_2 y))}}
										{\letin{y}{c_3}{(\pi_1 y, \inr(\pi_2 y))}}
									\right), \\
									l.\left(\caseof{\pi_1 l}{l_2.r_2}{l_3.r_3}
									\right)
								\end{array}}$
\item $\rsone{\ifthen {e_1}{e_2}{e_3}}{v}{q}$
\item $\reducetwo{q}{w}$
\end{enumerate}
By inversion of the splitting rule, we have
\begin{enumerate}
\item $\splitonesub 1 \rmbool$ and $\splitonesub 2 A$ and $\splitonesub 3 A$
\end{enumerate}
There are two rules for evaluation of an if.  We proceed with the true case first:
\begin{enumerate}
\item $\rsone{e_2}{\bf true}{q_1}$ and (something about $q_1$) and $\rsone{e_2}{v}{q}$
\item $\reduce{c_1}{\valprod{\bf true}{u_1}}$, by inductive hypothesis for 1
\item $\reduce{[u_1/l_1]1_2}{w}$, by inductive hypothesis for 1
\item $\reduce{c}{\valprod{v}{u_2}}$, by inductive hypothesis for 2
\item $\reduce{[u_2/l_2]r_2}{w}$, by inductive hypothesis for 2
\item $\reduce{\letin{y}{c}{(\pi_1(\pi_1 y),\pi_2 y)}}{\valprod{v}{u}}$
\end{enumerate}
From this, it can be directly proven that
\begin{enumerate}
\item $\reduce{\letin{y}{c}{(\pi_1(\pi_1 y),\pi_2 y)}}{\valprod{v}{u}}$
\item $\reduce{[u/l](\pi_1~r)}{w}$
\end{enumerate}
Which is what we wanted to show.

>>>>

\begin{figure}
\caption{\corelang~Diagonal Semantics}
\label{fig:refSemantics}
\begin{mathpar}
\inferRedsplitone [\rmunit]		{\red {()}{()}{()}}												{\cdot} 												\and
\inferRedsplitone [int]			{\red {i}{i}{()}}													{\cdot} 												\and
\inferRedsplitone [bool]			{\red {b}{b}{()}}													{\cdot} 												\and
\inferRedsplitone [hyp]			{\red {[u,x]}{u}{x}}												{\cdot}													\and
\inferSpeculation [\rmunit]		{\red {()}{()}}													{\cdot} 												\and
\inferSpeculation [int]			{\red {i}{i}}														{\cdot} 												\and
\inferSpeculation [bool]			{\red {b}{b}}														{\cdot} 												\and
\inferSpeculation [hyp]			{\red {x}{x}}														{\cdot}													\and
\inferRedsplitone [let]			{\red {\letin{x}{e_1}{e_2}}{u_2}{\letin {x}{r_1}{r_2}}}		{\sub [1] & \red{[[u_1,x]/x]e_2}{u_2}{r_2}} 		\and
\inferSpeculation [let]			{\red {\letin{x}{e_1}{e_2}}{\letin{x}{r_1}{r_2}}}				{\sub [1] & \sub [2]}									\and
\inferRedsplitone [\times I]		{\red {(e_1,e_2)}{\valprod{u_1}{u_2}}{\valprod{r_1}{r_2}}}	{\sub [1] & \sub [2]}									\and
\inferRedsplitone [\times E_1]	{\red {\pi_1~e}{u_1}{\pi_1 r}}									{\red {e}{\valprod{u_1}{u_2}}{r}}					\and
\inferRedsplitone [\times E_2]	{\red {\pi_2~e}{u_2}{\pi_2 r}}									{\red {e}{\valprod{u_1}{u_2}}{r}} 					\and
\inferSpeculation [\times I]		{\red {(e_1,e_2)}{(r_1,r_2)}}									{\sub [1] & \sub [2]}									\and
\inferSpeculation [\times E_1]	{\red {\pi_1~e}{\pi_1 r}}										{\red {e}{r}}											\and
\inferSpeculation [\times E_2]	{\red {\pi_2~e}{\pi_2 r}}										{\red {e}{r}} 											\and
\inferRedsplitone [if_T] 		{\red {\ifthen{e_1}{e_2}{e_3}}{u}{(r_1;r_2)}}					{\red{e_1}{\bf true }{r_1} & \red{e_2}{u}{r_2}}	\and
\inferRedsplitone [if_F] 		{\red {\ifthen{e_1}{e_2}{e_3}}{u}{(r_1;r_3)}}					{\red{e_1}{\bf false}{r_1} & \red{e_3}{u}{r_3}}	\and
\inferSpeculation [if] 			{\red {\ifthen{e_1}{e_2}{e_3}}{\ifthen{r_1}{r_2}{r_3}}}		{\sub [1] & \sub [2] & \sub [3]}					\\
\inferRedsplitone [\fut I]		{\red {\next~e}{()}{r}}											{\rstwosub}											\and
\inferSpeculation [\fut E]		{\red {\prev~e}{r}}												{\rsone {e}{()}{r}}
\end{mathpar}
\end{figure}

@@

\begin{figure*}
\caption{Term Splitting}
\label{fig:termSplit}
\newcommand {\projbind} [1] {\letin{l_{#1}}{\pi_{#1} l}{r_{#1}}}
\begin{mathpar}
\infersplitone [unit]			{\spl {()}{\rmunit}{((),()),\_.()}}															{\cdot} \and
\infersplittwo [unit]			{\spl {()}{\rmunit}{(),\_.()}}																{\cdot} \and
\infersplitone [int]			{\spl {i}{\rm int}{(i,()),\_.()}}															{\cdot} \and
\infersplittwo [int]			{\spl {i}{\rm int}{(),\_.i}}																	{\cdot} \and
\infersplitone [\times I] 	{\spl {(e_1,e_2)}{A\times B}
									{ \begin{array}{l}
									\left(
										\letin{y_1}{c_1}{\letin{y_2}{c_2}{
										((\pi_1 y_1, \pi_1 y_2),(\pi_2 y_1, \pi_2 y_2))}}
									\right), \\
									l.(\projbind 1,\projbind 2)
									\end{array}}}																				{\sub 1 A & \sub 2 B} \and
\infersplittwo [\times I]	{\spl {(e_1,e_2)}{A\times B}{(p_1,p_2),l.(\projbind 1,\projbind 2)}}					{\sub 1 A & \sub 2 B} \and
\infersplitone [\times E_1]	{\spl {\pi_1~e}{A}{\letin{y}{c}{(\pi_1(\pi_1 y),\pi_2 y)},l.\pi_1~r}}					{\sub {} {A\times B}} \and
\infersplitone [\times E_2]	{\spl {\pi_2~e}{B}{\letin{y}{c}{(\pi_2(\pi_1 y),\pi_2 y)},l.\pi_2~r}}					{\sub {} {A\times B}} \and
\infersplittwo [\times E_1]	{\spl {\pi_1~e}{A}{p,l.\pi_1~r}}															{\sub {} {A\times B}} \and
\infersplittwo [\times E_2]	{\spl {\pi_2~e}{B}{p,l.\pi_2~r}}															{\sub {} {A\times B}} \and
\infersplitone [hyp]			{\spl {v}{A}{(x,()),\_.()}}																	{x:A^\bbone \in \Gamma} \and
\infersplittwo [hyp]			{\spl {v}{A}{(),\_.x}}																		{x:A^\bbtwo \in \Gamma} \\
\infersplitone [let]			{\spl {\letin{x}{e_1}{e_2}}{B}{\cdots}}														{\sub 1 A & \sub [\Gamma,x:A^\bbone] 2 B} \and
\infersplittwo [let]			{\spl {\letin{x}{e_1}{e_2}}{B}{(p_1,p_2), l.\letin{x}{(\projbind 1)}{\projbind 2}}}	{\sub 1 A & \sub [\Gamma,x:A^\bbtwo] 2 B} \and
\infersplitone [\fut I]		{\spl {\next~e}{\fut A}{((),p),l.r}}														{\splittwosub {} A} \and
\infersplittwo [\fut E]		{\spl {\prev~e}{A} {\pi_2~c,l.r}}															{\splitonesub {} {\fut A}} \and
\infersplitone [if]			{\spl {\left( \tallif {e_1}{e_2}{e_3} \right)}{A} 
								{\begin{array}{l}
									\left(\tallif {\pi_1 c_1}
										{\letin{y}{c_2}{(\pi_1 y, \inl(\pi_2 y))}}
										{\letin{y}{c_3}{(\pi_1 y, \inr(\pi_2 y))}}
									\right), \\
									l.(\caseof{l}{l_2.r_2}{l_3.r_3})
								\end{array}}}																					{\sub 1 \rmbool & \sub 2 A & \sub 3 A } \and
\infersplittwo [if]			{\spl {\left(\tallif {e_1}{e_2}{e_3}\right)}{A} 
								{(p_1,p_2,p_3),l.\left(\tallif{(\projbind 1)}{(\projbind 2)}{(\projbind 3)}\right)}}	{\sub 1 \rmbool & \sub 2 A & \sub 3 A } \and
\end{mathpar}
\end{figure*}

@@

\item {\bf If.} Suppose that
\begin{enumerate}
\item $\splitone {\left( \tallif {e_1}{e_2}{e_3} \right)}{A} 
								{\begin{array}{l}
									\left(\tallif {\pi_1c_1}
										{\letin{y}{c_2}{(\pi_1 y, \inl(\pi_2 y))}}
										{\letin{y}{c_3}{(\pi_1 y, \inr(\pi_2 y))}}
									\right), \\
									l.\left(\caseof{l}{l_2.r_2}{l_3.r_3}
									\right)
								\end{array}}$
\item $\rsone{\ifthen {e_1}{e_2}{e_3}}{v}{q}$
\item $\reducetwo{q}{w}$
\end{enumerate}
By inversion of the splitting rule, we have
\begin{enumerate}
\item $\splitonesub 1 \rmbool$ and $\splitonesub 2 A$ and $\splitonesub 3 A$
\end{enumerate}
There are two rules for evaluation of an if.  We proceed by inverting the if-true and evaluation rules first:
\begin{enumerate}
\item $q$ has the form $(q_1;q_2)$
\item $\reducetwo{q_1}{w_1}$ and $\reducetwo{q_2}{w}$
\item $\rsone{e_1}{\bf true}{q_1}$ and $\rsone{e_2}{v}{q_2}$
\end{enumerate}
We can now invoke the inductive hypothesis for the predicate to yield
\begin{enumerate}
\item $\reduce{c_1}{\valprod{\bf true}{u_1}}$
\item $\reduce{[u_1/l_1]q_1}{w_1}$
\end{enumerate}
We can also invoke the inductive hypothesis for the branch to yield
\begin{enumerate}
\item $\reduce{c_2}{\valprod{v}{u_2}}$
\item $\reduce{[u_2/l_2]r_2}{w}$
\end{enumerate}
From this, it can be directly proven that
\begin{enumerate}
\item $\reduce{
		\left(\tallif {\pi_1c_1}
			{\letin{y}{c_2}{(\pi_1 y, \inl(\pi_2 y))}}
			{\letin{y}{c_3}{(\pi_1 y, \inr(\pi_2 y))}}
		\right)}{\valprod{v}{\inl~u_2}}$
\item $\reduce{[\inl~u_2/l](\caseof{l}{l_2.r_2}{l_3.r_3})}{w}$
\end{enumerate}
Everything works similarly for the if-false case.  Either way, we have what we want to show.
